\vspace{-1.5em} \bf A general and efficient representation of ancestral recombination graphs

### 2023-10-27 paper.tex

\begin{abstract}
New developments have made it possible to infer genetic genealogies in
the presence of recombination at scale, enabling many
downstream applications in population and statistical genetics.
The structure representing such recombinant genetic ancestry
is usually referred to as an ancestral recombination graph (ARG),
% FIXME this sentance needs to change. We want to capture more of the
% idea of the formalisms falling behind actual practise and the
% confusion this is causing
although there is some confusion about the interpretation and little
agreement on specific details.
We propose a concrete definition of ARGs
in terms of genomes and their intervals of genetic inheritance (gARG)
and contrast this with the classical event-based definition (eARG).
We show that eARGs are limited in the patterns
of genetic inheritance that can be represented,
and require a precise description of the details of all recombination events.
We show, in contrast, that the gARG definition
can fully capture the richness of modern large-scale datasets,
enables fine-grained
levels of precision about recombination to be represented,
and forms the basis of an efficient computational framework.
\end{abstract}

>>>>

\begin{abstract}
New developments have made it possible to infer genetic genealogies in
the presence of recombination at scale, enabling many
downstream applications in population and statistical genetics.
The structure representing such recombinant genetic ancestry
is usually referred to as an ancestral recombination graph (ARG).
% Be specific about what we're talking about.
ARGs are classically defined by the graph of
% We can play a bit loose here, please don't change to "common ancestor"
coalescence and recombination events in the history of sample,
with the differential inheritance structure encoded by the
crossover breakpoints associated with recombination events.
We show that this ``event ARG'' (eARG) encoding
is limited in the patterns of genetic inheritance that can be represented,
being insufficient to describe the outputs of modern ARG simulation
and inference methods,
and ill-suited to capturing the complex patterns of
relatedness that exist in contemporary datasets.
We propose an alternative ``genome ARG'' (gARG) encoding
expressed in terms of
% we haven't define a "genome" yet, so need to give a bit of context
specific genomes and their intervals of genetic inheritance.
We show it can fully capture the richness of modern large-scale datasets
and is compatible with modern simulation and inference methods.
Moreover, we show that, unlike the eARG encoding which requires
a precise estimate of all recombination events,
the gARG encoding enables fine-grained
levels of precision about recombination to be represented,
opening up a rich set of additional details.
\end{abstract}
